---
title: Investigating the Relationship Between Corruption Perception and Happiness
  Across Countries
author: 'Aayushi'
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  slidy_presentation:
    font_adjustment: -1
    highlight: haddock
    mathjax: default
    df_print: paged
  ioslides_presentation:
    highlight: haddock
  beamer_presentation:
    highlight: haddock
subtitle: An Applied Analytics Investigation Using World Happiness Data (2015–2020)
---

```{r setup, include=FALSE}
# load required pacakges

knitr::opts_chunk$set(
  fig.align = "center",
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  fig.width = 5,
  fig.height = 3
)

# Packages loaded 

library(ggplot2) 
library(dplyr)  
library(knitr) 

#load data

data <- read.csv("WorldHappiness_Corruption_2015_2020.csv")

```


# Introduction

- Happiness is influenced by social, economic, and political conditions across countries. 
- Corruption perception may affect public trust, quality of life, and overall wellbeing. 
- Countries with lower perceived corruption are often associated with stronger institutions and higher living standards. 
- This investigation analyses the relationship between Corruption Perception Index (CPI) scores and happiness scores using global data collected between 2015 and 2020.



# Problem Statement

- This investigation aims to examine whether perceived corruption is associated with happiness levels across countries. 
- The analysis focuses on the relationship between Corruption Perception Index (CPI) scores and happiness scores from 2015 to 2020. 
- Descriptive statistics and visualisations will be used to identify trends and distribution patterns in the data. 
- Hypothesis testing and linear regression analysis will be applied to evaluate the statistical relationship between corruption perception and happiness scores



# Data

- The dataset used in this investigation contains World Happiness and Corruption Perception Index (CPI) data collected between 2015 and 2020. 
- The data includes 792 observations and 13 variables related to happiness, economic conditions, freedom, government trust, and corruption perception across countries. 
- The dataset was obtained from Kaggle and combines information from global happiness reports and CPI data sources. 
- The main variables used in this investigation are: 
     - `happiness_score`: overall happiness score of a country 
     - `cpi_score`: Corruption Perception Index score (higher values indicate lower perceived corruption) 
     - `gdp_per_capita`: economic indicator representing income level 
     - `continent`: categorical variable identifying geographical region 
- Only variables relevant to the investigation were selected for analysis.



# Data Cont.

- Data preprocessing was performed before statistical analysis. 
- Only variables relevant to the investigation were selected to improve clarity and reduce unnecessary complexity. 
- Missing values were checked using R functions and handled appropriately during the analysis process. 
- Numerical variables such as `happiness_score`, `cpi_score`, and `gdp_per_capita` were retained for statistical modelling. 
- The variable `continent` was used as a categorical grouping variable for comparisons and visualisation. 
- The dataset was analysed using R statistical software and relevant visualisation packages.



# Descriptive Statistics and Visualisation

- Descriptive statistics were used to summarise the distribution and variability of the main variables.
- The variables analysed include `happiness_score`, `cpi_score`, and `gdp_per_capita`. 
- Measures such as mean, median, standard deviation, minimum, and maximum values were examined. 
- Visualisations were used to identify patterns, skewness, trends, and possible outliers within the data.

```{r, echo=FALSE}
table1 <- data.frame(
  Variable = c("Happiness Score", "CPI Score"),
  Mean = c(5.47, 44.33),
  Median = c(5.49, 38.00),
  SD = c(1.12, 19.51),
  Min = c(2.57, 11.00),
  Max = c(7.81, 91.00)
)

knitr::kable(
  table1,
  digits = 2,
  caption = "Table 1: Summary Statistics"
)
```



# Descriptive Statistics Cont.

- Table 1 shows that the average happiness score across countries was 5.47, with a median of 5.49.
- The mean and median happiness scores are very similar, suggesting that the distribution of happiness
scores is approximately symmetric.
- The average CPI score was 44.33, while the median CPI score was 38.00.
- The difference between the mean and median CPI scores suggests a degree of positive skewness in the CPI distribution.
- The standard deviation of happiness scores was 1.12, indicating moderate variability, while the CPI score showed greater variability with a standard deviation of 19.51.



# Relationship between CPI and Happiness

```{r, echo=FALSE, fig.width=6, fig.height=4}
ggplot(data, aes(x = cpi_score, y = happiness_score)) +
  geom_point(color = "steelblue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Figure 1: Relationship Between CPI Score and Happiness Score",
    x = "CPI Score",
    y = "Happiness Score"
  ) +
  theme_minimal()
```



# Scatterplot Interpretation

- Figure 1 shows a positive relationship between CPI scores and happiness scores.
- Countries with higher CPI scores, meaning lower perceived corruption, generally report higher happiness levels.
- The upward trend line suggests a positive association between corruption perception and happiness.
- However, some variation remains around the trend line, indicating that happiness is also influenced by other social and economic factors.




# Distribution of Happiness Scores

```{r, echo=FALSE}
ggplot(data, aes(x = happiness_score)) +
  geom_histogram(bins = 20, fill = "skyblue", color = "black") +
  labs(
    title = "Figure 2: Distribution of Happiness Scores",
    x = "Happiness Score",
    y = "Frequency"
  ) +
  theme_minimal()
```



# Distribution Interpretation

- Figure 2 displays the distribution of happiness scores across countries.
- The distribution appears approximately symmetric.
- The mean happiness score (5.47) and median happiness score (5.49) are very similar, indicating little evidence of strong skewness.
- The distribution suggests that the data are reasonably suitable for further statistical analysis and modelling.




# Hypothesis Testing

- - A simple linear regression model was fitted to investigate the relationship between CPI score and happiness score:

```{r, echo=FALSE}
model1 <- lm(happiness_score ~ cpi_score, data = data)

reg_table <- data.frame(
  Statistic = c("Slope Coefficient", "p-value", "R-squared"),
  Value = c("0.03995", "<0.001", "0.4803")
)

knitr::kable(reg_table)
```

- The following hypotheses were tested:

$$H_0: \beta_1 = 0$$

$$H_A: \beta_1 \neq 0$$

- Where $\beta_1$ represents the slope of the relationship between CPI score and happiness score.




# Hypothesis Testing Cont.

- The regression model found a statistically significant positive relationship between CPI score and happiness score (p < 0.001).
- Since the p-value is less than the significance level of 0.05, the null hypothesis was rejected.
- The estimated slope coefficient was 0.03995, indicating that for every one-unit increase in CPI score, the predicted happiness score increases by approximately 0.04 points.
- The model explained approximately 48.03% of the variation in happiness scores (R² = 0.4803).
- These findings suggest that countries with lower perceived corruption tend to report higher levels of happiness.



# Model Diagnostics

```{r, echo=FALSE, fig.width=7, fig.height=5}
par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
plot(model1)
```



# Model Diagnostics Cont.

- Diagnostic plots were examined to assess the assumptions of the linear regression model. 
- The Residuals vs Fitted plot showed a generally random pattern around zero, suggesting that the assumption of linearity was reasonably satisfied.
- The Normal Q-Q plot indicated that residuals were approximately normally distributed, with only minor deviations observed in the tails.
- The Scale-Location plot suggested relatively constant variance of residuals across fitted values, indicating no serious heteroscedasticity.
- The Residuals vs Leverage plot did not identify any highly influential observations that would substantially affect the model results.
- Overall, the diagnostic plots suggest that the assumptions of linear regression were reasonably satisfied and the model was appropriate for this investigation.



# Discussion

- The results indicate a positive relationship between CPI score and happiness score across countries.
- Countries with lower perceived corruption generally reported higher happiness levels.
- The regression model explained approximately 48% of the variation in happiness scores, suggesting that corruption perception is an important factor but not the only determinant of happiness.
- A strength of this investigation is the use of data from multiple countries across several years.
- A limitation is that other factors such as income, health, and social support may also influence happiness but were not included in the model.
- Future investigations could examine multiple factors simultaneously using multiple regression analysis.



# Conclusion

- This investigation examined the relationship between corruption perception and happiness across countries.
- Descriptive statistics and visualisations suggested a positive association between CPI score and happiness score.
- Linear regression analysis found a statistically significant positive relationship between CPI score and happiness score (p < 0.001).
- The model explained approximately 48% of the variation in happiness scores.
- Overall, the findings suggest that countries with lower perceived corruption tend to report higher levels of happiness.



# References

- World Happiness Report. (2024). World Happiness Data.https://www.worldhappiness.report/
- Transparency International. (2024). Corruption Perceptions Index (CPI).https://www.transparency.org
- Kaggle. World Happiness and Corruption Dataset.https://www.kaggle.com
- Wickham, H. et al. (2023). ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics.
- Wickham, H. et al. (2023). dplyr: A Grammar of Data Manipulation.